---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zmij

<!-- badges: start -->
[![R-CMD-check](https://github.com/nanxstats/zmij/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nanxstats/zmij/actions/workflows/R-CMD-check.yaml)
![CRAN status](https://www.r-pkg.org/badges/version/zmij)
[![extendr](https://img.shields.io/badge/extendr-^0.9.0-276DC2)](https://extendr.github.io/extendr/extendr_api/)
<!-- badges: end -->

zmij converts double-precision floating-point values to decimal strings with
the shortest significands needed for round-trip recovery. It provides a
vectorized R interface to the [zmij](https://github.com/dtolnay/zmij) Rust
crate, a port of Victor Zverovich's
[zmij algorithm](https://github.com/vitaut/zmij).

## Installation

You can install the development version of zmij from GitHub with:

``` r
# install.packages("pak")
pak::pak("nanxstats/zmij")
```

## Usage

Format a numeric vector:

```{r example}
library(zmij)

x <- c(pi, 0.1, -0, .Machine$double.xmin, .Machine$double.xmax)
text <- format_double(x)
text
identical(parse_double(text), x)
```

Missing and non-finite values are handled explicitly:

```{r special-values}
format_double(c(NA_real_, NaN, Inf, -Inf))
```

`parse_double()` is also useful directly when correctly rounded conversion is
important, especially near the limits of the binary64 range:

```{r parse}
parse_double(c("1.5", "3e-1", "5e-324", "1.7976931348623157e+308"))
```

Names and dimensions are preserved:

```{r shape}
format_double(matrix(c(0.1, pi, 1e-6, 1e16), nrow = 2))
```
